---
title: "README"
author: "Ingo Rohlfing"
date: "`r Sys.Date()`"
output: 
    github_document: default
---

```{r, include = FALSE}
knitr::opts_chunk$set(eval = F)
```


## Short explainer on motivation
R Markdown reports are a great tool for turning dynamic code and text into a 
full-fledged report as PDF, Word file or another format. When one is familiar
with R Markdown reports, then one most likely knows Rmd scripts visually
separating text from chunks with code in them. 

What I only learned over time, slowly, about R Markdown, are two related 
things: 

1. One can transform Rmd files into R scripts and vice versa.
2. One can render R scripts into R Markdown reports.

As probably everything with R Markdown, this is well documented in the awesome
[R Markdown Cookbook](https://bookdown.org/yihui/rmarkdown-cookbook/)
by Xie, Dervieux and Riederer. Still, I wanted to briefly address this here 
for my own purposes and because others might find this then useful too.

## From Rmd to an R script
When you have an Rmd file like [this one](Rmd_to_R_script.Rmd) and want to 
turn it into an R script, you need `knitr::purl()`. For the same Rmd file,
you can get three R scripts with different levels of documentation. For the
argument `documentation = 0`, one gets [a plain R script](just_script_0.R) 
out of the Rmd file.
```{r}
knitr::purl(input = "Rmd_to_R_script.Rmd", 
            output = "just_script_0.R",
            documentation = 0) # give me just the code
```

