---
title: "README"
author: "Ingo Rohlfing"
date: "`r Sys.Date()`"
output: 
    github_document: default
---

```{r, include = FALSE}
knitr::opts_chunk$set(eval = F)
```


## Short explanation of motivation
R Markdown reports are a great tool for turning dynamic code and text into a 
full-fledged report as PDF, Word file or another format. Everyone who is 
familiar with R Markdown reports most likely knows Rmd scripts that visually 
separate text and code output from chunks containing code. 

There are two related facts about R Markdown that I only learned slowly, 
over time:  

1. One can transform Rmd files into R scripts and vice versa.
2. One can render R scripts into R Markdown reports.

As is probably true about all aspects of R Markdown, this is well documented 
in the awesome 
[R Markdown Cookbook](https://bookdown.org/yihui/rmarkdown-cookbook/)
by Xie, Dervieux and Riederer. Still, I wanted to post this here because I

- wanted to go through the commands once myself and convert actual scripts;
- believe these very useful features of R Markdown are not as widely known as
they probably should.

## From an Rmd file to an R script
When you have an Rmd file like [this one](Rmd_to_R_script.Rmd) and want to 
turn it into an R script, you need `knitr::purl()`. You can convert an Rmd file
into an R script in *three different ways*, depending on how much information
you want to preserve. This can be specified with the `documentation` argument
where:

- `0`: plain code + annotations;
- `1`: plain code + annotations + chunk information;
- `2`: the same information as in Rmd file, just in a different format.

### Rmd => plain code
This command reduces [the Rmd file](Rmd_to_R_script.Rmd) to [an R script only 
with code and annotations](just_code_0.R). Since the identifying information is 
part of the Rmd header, this information is now gone, as is all text.
```{r, eval = F}
knitr::purl(input = "Rmd_to_R_script.Rmd", 
            output = "just_code_0.R",
            documentation = 0) # just the code
```

For illustration, this part of the Rmd file...

````
Load some toy data.
```{r, demo-chunk, echo = T}`r ''`
# load some data
data(mtcars)
```
````

becomes this in the script.
```
# load some data
data(mtcars)
```

### Rmd => code + chunk information
This command reduces [the Rmd file](Rmd_to_R_script.Rmd) to [an R script only 
with code, annotations and chunk info](code_chunkinfo_1.R). 
```{r, eval = F}
knitr::purl(input = "Rmd_to_R_script.Rmd", 
            output = "code_chunkinfo_1.R",
            documentation = 1) # code + chunk information
```

For illustration, this part of the Rmd file...

````
Load some toy data.
```{r, demo-chunk, echo = T}`r ''`
# load some data
data(mtcars)
```
````

becomes this in the script.
```
## ----demo-chunk, echo = T----------------------------------------------------
# load some data
data(mtcars)
```

When you turn the two R scripts into reports, they almost look the same because
the content is the same (compare the PDF reports for 
[plain code](just_code_0.pdf) to [chunk info + code](code_chunkinfo_1.pdf). 
The only reason I see for converting with `documentation = 1` is that you can 
directly navigate to a selected chunk, as well as easily moving back and forth
between them.

### Rmd => identical R script
This command converts [the Rmd file](Rmd_to_R_script.Rmd) to 
[an R script that has exactly the same content](complete_script_2.R). 
```{r, eval = F}
knitr::purl(input = "Rmd_to_R_script.Rmd", 
            output = "complete_script_2.R",
            documentation = 2) # full conversion
```

For illustration, this part of the Rmd file...

````
Load some toy data.
```{r, demo-chunk, echo = T}`r ''`
# load some data
data(mtcars)
```
````

becomes this in the script.
```
#' Load some toy data.
## ----demo-chunk, echo = T----------------------------------------------------
# load some data
data(mtcars)
```

The formatting is different in the script, but the exact same information is 
preserved. You will see the knitted PDFs are the same when you prepare a PDF 
report for [the Rmd file](Rmd_to_PDF.pdf) and the 
[converted R script](complete_script_2.pdf). (I haven't edited the R script
after conversion, so they also have the same title.)

## From an R script to an Rmd file
When you have an R script like [this one](R_script_to_Rmd.R) and want to 
turn it into an Rmd file, you need `knitr::spin()`. This is the command doing
the job.
```{r spin}
knitr::spin(hair = "R_script_to_Rmd.R", 
            format = "Rmd", 
            knit = FALSE) # we don't an html and md file for the R script now
```

For illustration, this part of the R script...
```
#' Load some toy data.
#+ demo-chunk, echo = TRUE
# load some data
data(mtcars)
```

becomes this in the Rmd file.
````
Load some toy data.

```{r demo-chunk, echo = TRUE}`r ''`
# load some data
data(mtcars)
```
````

When you look at the [generated Rmd file](R_script_to_Rmd.Rmd), it looks 
exactly like the Rmd file `Rmd_to_R_script.Rmd` that was previously converted to
R scripts (except for the different title in the file). You only need to follow 
a small number of formatting rules when you want a nice Rmd file from 
`knitr::spin()`, which you can see in the [R script](R_script_to_Rmd.R) used as
input (here is information about formatting in the
[R Markdown cookbook](https://bookdown.org/yihui/rmarkdown-cookbook/spin.html)). 

## Turn Rmd and R scripts into a report
Rmd files are known for allowing one to simply create reports (in html format,
PDF etc.) by hitting the "Knit" button in RStudio. The command line 
equivalent is `rmarkdown::render()`, which, at a minimum, only needs the 
Rmd file as input. For example, you can create the 
PDF [RMD_to_pdf](Rmd_to_PDF.pdf) with the following line. PDF as an output
format is specified in the Rmd header, so I omit the `output_format` argument
here.
```{r RMD-to-PDF}
rmarkdown::render(input = "Rmd_to_R_script.Rmd", 
                  output_file = "Rmd_to_PDF")
```

The good thing now is that `rmarkdown::render()` works equally well with *any*
R script regardless of how it has been formatted. Of course, it makes more sense
to render a file such as [R_script_to_Rmd.R](R_script_to_Rmd.R) because it 
produces a more attractive PDF. [This PDF](R_script_to_PDF.pdf) has been
produced with this command.
```{r R-script-to-PDF}
rmarkdown::render(input = "R_script_to_RMD.R", 
                  output_file = "R_script_to_PDF") # PDF is specified in header
```


Another good feature of `rmarkdown::render()` is that you can create a report
without having the file that is to be rendered open in RStudio.

### So, why is all of this of interest?
I think it is good to know about the two-way street between Rmd files and R 
scripts and the all-purpose use of `rmarkdown::render()`. The main advantage is 
that one does not have to think much about what format to start with in an 
analysis. Rmd files and R scripts have both advantages and disadvantages and 
one format is not clearly superior to the other.  
In some of my courses, for example, some course participants found it
easier to work with Rmd files, while others could better work with R scripts.
The good news about `purl()`, `spin()` and `render()` is that everyone can
work with the preferred format, produce the same beautiful reports and easily
convert Rmd files to R scripts and vice versa, if necessary.

### Packages used (on the surface)
Hadley Wickham, Jim Hester and Winston Chang (2021). *devtools: Tools to Make
Developing R Packages Easier.* R package version 2.4.2.
<https://CRAN.R-project.org/package=devtools>

JJ Allaire and Yihui Xie and Jonathan McPherson and Javier Luraschi and Kevin 
Ushey and Aron Atkins and Hadley Wickham and Joe Cheng and Winston Chang and 
Richard Iannone (2021). *rmarkdown: Dynamic Documents for R.*
R package version 2.11. <https://rmarkdown.rstudio.com>.

Yihui Xie (2021). *knitr: A General-Purpose Package for Dynamic Report 
Generation in R.* R package version 1.36.

Yihui Xie (2015). *Dynamic Documents with R and knitr.* (2nd ed.) 
Chapman and Hall/CRC. 

Yihui Xie (2014). *knitr: A Comprehensive Tool for Reproducible Research in R.*
In Victoria Stodden, Friedrich Leisch and Roger D. Peng, editors, 
*Implementing Reproducible Computational Research*. Chapman and Hall/CRC. 

Yihui Xie and J.J. Allaire and Garrett Grolemund (2018). *R Markdown: 
The Definitive Guide.* Chapman and Hall/CRC. 
<https://bookdown.org/yihui/rmarkdown>.

Yihui Xie and Christophe Dervieux and Emily Riederer (2020). 
*R Markdown Cookbook.* Chapman and Hall/CRC. 
<https://bookdown.org/yihui/rmarkdown-cookbook>.

### Session parameters
```{r, echo = FALSE, eval = TRUE}
devtools::session_info()
```


